Title here

### 2016-12-06 portfolio/text/ms.Rmd
weird sync issue

```{r set-knitr-options, cache=FALSE, echo=FALSE}
library("knitr")
opts_chunk$set(cache=TRUE, echo=FALSE, message=FALSE, autodep=TRUE,
  results='hide')
opts_chunk$set(warning=FALSE, message=FALSE, cache.path="../cache/",
  cache.lazy=TRUE)
```
```{r get-dat, cache=FALSE}
load("../data-generated/output-values.rda")
h1 <- readRDS("../data-generated/h1_effect.rds")
spoke <- readRDS("../data-generated/spoke_fraction.rds")
g2_dens <- readRDS("../data-generated/g2_prob_dens_above_zero.rds")
n_grouped_permits <- readRDS("../data-generated/n_grouped_permits.rds")
```

# Abstract

A fundamental challenge for individuals who rely on the extraction of natural
resources for their livelihood is income variability. Fishers, loggers, farmers,
and miners, among others, face high levels of income uncertainty driven by a mix
of environmental, biological, and economic variability. A key component
underlying the sustainability of these industries is identifying both how
regulatory actions and individual behaviour affect income variability and by
extension the economic stability and sustainable use of natural resources. For
commercial fisheries, recent work has identified that targeting a diversity of
species can reduce income variability. However, any diversity-variability
relationship is embedded in a regulatory framework that restricts individual
ability to change fishing behaviour. Here we identify three primary axes by
which fishers in Alaska, one of the world's most productive fishery regions, can
potentially diversify fishing activities and investigate how individual actions
along each diversity axis affect both revenue and revenue variability. We
construct a novel statistical model to simultaneously assess the value to
individual fishers of choosing to (1) own permits that allow targeting a diverse
fish species assemblage, (2) diversify by owning multiple permits, or (3) catch
a greater species diversity within the constraints of a given permit. We show
contrasting effects of diversity on revenue and revenue variability along the
three axes. First, we show that fishing permits with higher species diversity
reduces variability (`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction
per unit increase in mean effective species diversity). Next, we show that
fishing an additional permit is associated with higher revenue and lower
variability (e.g., `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`%
IQR] lower expected standard deviation for halibut plus another permit vs.
single halibut permits). Finally, we show that increasing species diversity
within the constraints of a permit set has a fishery-dependent effect on revenue
and is usually (`r g2_dens`% probability) associated with increased uncertainty
in revenue the following year. Thus, while there are clear benefits to some axes
of diversification, others are less effective. Importantly, the most effective
measure for increasing revenue and decreasing risk --- purchasing additional
permits to participate in additional fisheries --- was also the most costly. At
the same time, Alaskan commercial fisheries have been increasingly focussed on
single-permit (often salmon) fishing strategies. Our results suggest that the
benefit of species diversity for revenue variability is mostly an across-permit
effect, not a benefit that can be easily harnessed by an individual permit
holder. Effective methods of diversification --- purchasing additional permits
or switching permits --- are costly and often limited by regulations such as
limited entry programs and individual transferable quotas (ITQs). Take home ...

\bigskip
\noindent

\clearpage

<!-- Extra: -->

<!-- - there is often limited room for intentional diversification within a permit -->
<!-- - in fact, specializing can be profitable within some fisheries -->
<!-- - Eric: analog here in farming literature is that diversification -->
<!--   for small scale farmers generally good, except when they -->
<!--   specialize on high demand/high value crops (tobacco, cocoa). -->
<!--   Seems like benefits of diversification are also smaller for -->
<!--   large-scale industrial farmers. -->

<!-- \noindent -->
<!-- Main message: -->
<!-- Diversified portfolios of fishing permits and species reduce risk for fishermen, -->
<!-- but changing species diversity can be risky, costly, and limited by regulations. -->

<!-- , which is concerning  -->
<!-- given that some fishing regulations and market forces -->
<!-- may be increasingly promoting specialization. -->
<!-- Main message in 25 words or less: -->

<!-- Take 1: -->
<!-- - There is a trend towards specializing on single salmon permits in Alaska -->
<!-- - Diversified portfolios of fishing permits and species reduce risk, -->
<!--   but changing species diversity for an individual permit holder can be risky -->
<!-- - The types of diversity needed to reduce risk are often restricted by -->
<!--   cost and regulations -->

<!-- Take 2: -->
<!-- - The effect of species diversification on revenue and risk for fishermen -->
<!-- depends on the type of diversification and the fishery. -->

<!-- Goal: Introduction, Methods, Results, Discussion, Conclusion:
<!-- 600, 900, 500, 800, 200, -->

# Introduction

<!-- Papers to cite [@smith1986] -->

Managing natural resources and sustaining the livelihoods of those that depend
on their extraction is difficult. These professions tend to have high
year-to-year variability in income [@mishra2002; @sethi2012b]. In addition to
economic risk, such as changes in demand, individuals dependent on natural
resources are also subject to biological and environmental variability. For
example, drought and flooding are a major source of risk to agricultural food
security and farmers' incomes [@gregory2005], and catastrophic disease outbreaks
and wildfires increase risk for the logging industry [@kirilenko2007].

Individuals who rely on natural resources for income develop strategies to
reduce income variability. For example, farmers may rotate their crops to buffer
short-term climate variation and long-term environmental trends [@zentner1988].
Similarly, loggers maintain a diverse mix of tree species, age structure, and
harvest in geographical areas to buffer against disease outbreak, wildfires, and
price variation REF. However, otherwise well-intentioned regulations can limit
the actions available to individuals to reduce income variability, or
incentivize against diversification. For example, in terrestrial agriculture,
crop subsidies in the US may incentivize some farms to specialize (e.g. dairy,
grains, soybeans; @mishra2002). In other instances, external market forces
(prices, demand) may discourage the adoption of diverse strategies and help
promote specialization [@bradshaw2004].

Commercial fishers present an interesting example of diversification, given the
extreme risk that some individuals experience, particularly relative to
terrestrial agriculture [@sethi2012b]. The majority of insurance programs
available to terrestrial farmers (e.g. crop insurance programs) are not
available to commercial fishers [@kasperski2013] and both fish and fishers are
spatially mobile. However, fishers still have the ability to lower risk. They
can, for example, target multiple species as permitted by a license, own
multiple licenses, or diversify spatially and fish in multiple regions. The
effectiveness of these diversification strategies for improving revenue and
reducing revenue variability remains uncertain. Furthermore, theory and
empirical studies both suggest there can be substantial benefits to
specialization (REFS). The majority of papers that examine effects of
diversification on revenue do so at aggregate levels (e.g. households, counties,
communities; REFS), rather than focus on individuals. Because of this, key
uncertainties remain about the fundamental relationship between diversity,
fishing revenue, and fishing revenue variability at the level of the individual.

The state and federal fisheries of Alaska provide a unique opportunity to
understand the role of individuals' strategies in determining income and income
variability for those whose livelihood depends on the extraction of natural
resources. Fisheries in Alaska are among the highest-valued fisheries in the
world with 2.7 billion kg landed biomass valued at nearly 2 billion USD annually
[@nmfs2016]. The industry supports a range of fishing fleets ranging from small,
open, single-operator vessels to multi-million-dollar state-of-the-art ships
with large crews. Like much of the rest of the world [@hilborn2003a], the days
of open, unregulated fisheries in most developed countries have passed ---
fishers must own a license to participate in virtually all fisheries. And while
such regulations have proven helpful in reducing or eliminating overfishing,
they limit the available strategies to reduce income variability. For example,
many permits limit fishing to a single species, diversifying by purchasing
additional fishing permits can be expensive (e.g. ~\$XX for king crab), and some
fisheries limit the entry of new participants through individual fishing quotas
(IFQs) (REF).

Using a unique dataset collected over 30 years (1985--2014) on individuals,
rather than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals rather than
vessels we can track people that switch fishery sectors over time, or alter
strategies and use multiple vessels. We develop and apply a hierarchical
Bayesian model to understand the effect of diversity on both individual revenue
and revenue variability that builds on results from vessel [@kasperski2013],
fleet [@sethi2012b], and community-level variability [@sethi2014]. Our model
controls for fishing effort and latent time effects (e.g. changing environmental
conditions), simultaneously estimates the effect of species diversity on
individual revenue and revenue variability, and estimates effects both within
and across fishing strategies (sets of similar licenses or permits). To our
knowledge, this is the first application of this type of model in any natural
resource field.

\clearpage

# Methods

We developed a hierarchical variance-function regression model, which separately
models changes in the mean and variance of individual revenue as a function of
covariates [@davidian1987; @hall1989]. Whereas traditional hierarchical linear
models assume homoscedasticity --- that the residual variance does not vary
systematically --- the variance-function regression model allows the variance to
be heteroscedastic and vary as a function of covariates. Our specific model
allows certain fishing strategies and levels of species diversification to
generate more or less revenue and be more or less variable than others.

## Data

We obtained commercial fisheries landings and revenue data for all permit
holders in Alaska from `r range_yrs[1]` to `r range_yrs[2]` from the Alaskan
Commercial Fisheries Entry Commission (CFEC). These data represent a total of
nearly `r round(ndat$initial_tot_rev/1e9)` billion 2009 USD and a combined `r 
round(ndat$initial/1e6,1)` million rows of data. We adjusted for inflation by
setting all revenues to 2009 USD [@bea2016]. We implemented a number of
filtering steps to focus on individuals actively engaged in commercial fishing
(Supporting Information). In particular, we removed permit holders whose median
revenue was less than \$10,000 2009 USD to focus on individuals for whom
commercial fishing formed a substantial part of their income.

We combined similar permits into permit groupings and identified sets of permit
groupings held by an individual in a specific year as "strategies" (Supporting
Information). For example, we combined permits that were the same except for
vessel size (e.g. a longliner fishing sablefish on a vessel under 60 ft was
considered to have the same strategy as a longliner fishing sablefish on
a vessel over 60 ft), and we combined a number of permits that were targeting
a single species and differed only in region fished. Therefore, for example,
someone fishing herring roe in Kodiak and someone fishing herring roe in Alaska
Peninsula were both considered to have a 'herring roe' permit group. For salmon
permits, we grouped permits using the same gear type across regions for groups
that shared similar species composition. After this process of combining
permits, we were left with `r n_grouped_permits` unique fishing permit groups
(Table S\ref{tab:strategies}), which we aggregated within each person-year
combination to form permit strategies. For example, if someone fished halibut
and sablefish in the same year, their strategy for that year would be
"halibut-sablefish". Specific details on our strategy definitions are available
in the Supporting Information.

Our models focused on species diversity as a possible predictor of revenue and
revenue variability while controlling for effort. For consistency with previous
analyses [@kasperski2013], we calculated effective species diversity for
permit holder $i$ and time (year) $t$ as the inverse of Simpson's diversity
index $\lambda$ [@simpson1949; @jost2006] (also referred to in economics as the
Herfindahl-Hirschman Index) weighted by revenue $R$:

\begin{equation}
1/\lambda_{i,t} = 1 / \left(\sum\limits_{s=1}^{n_{i,t}}
  \left(R_{i,s,t}/ \sum\limits_{s=1}^{n_{i,t}} R_{i,s,t}\right)\right)^2,
\end{equation}

\noindent where $s$ indexes species from $1$ through $n$. We calculated effort
or effective season length as the sum of the season lengths for each permit each
year. Season length for a given permit was calculated as the number of days
between the first instance of fishing a permit and the last instance of fishing
a permit each year. For example, if someone fished for 30 days on a salmon drift
gillnet permit in Southeast Alaska and 40 days on a statewide longline halibut
permit, that individual's season length would be 70 days.

## Hierarchical model

To jointly model revenue and revenue variability as a function of
individual-level covariates, we extended the basic variance-function regression
to a hierarchical Bayesian variance-function regression model with Gaussian
errors predicting $\log$ revenue, $\log(R_{i,t})$, for each fisher $i$ and time
$t$:

\begin{equation}
\log(R_{i,t}) \sim \mathrm{N} \left(\widehat{\log(R_{i,t})}, \widehat{\sigma_{i,t}^2} \right).
\end{equation}

\noindent Each year $t$ of revenue for fisher $i$ is assigned a strategy indexed
by $j$ (Supporting Information, Table \ref{tab:strategies}) and expected
$\log(R_{i,t})$ is modelled as:

<!-- Ole: I might swap out log(R)-hat for something like mu(or your favorit greek letter) just to avoid -->
<!-- potential confusion like using the term "expected log(R)".   -->
<!-- I also just don't like hats... but that's my problem. -->

\begin{equation}
\label{eq:mean}
\widehat{\log(R_{i,t})} = \beta_{0,j,t} -
  \beta_{1,j} S_{i,t} I_{i,t} +
  \beta_{2,j} S_{i,t} I_{i,t} +
  \beta_3 D_{i,t} -
  \beta_4 S_{i,t} D_{i,t} I_{i,t} +
  \beta_5 S_{i,t} D_{i,t} I_{i,t} +
  \log(R_{i,t-1}),
\end{equation}

\noindent where $S_{i,t}$ is the percent change in species diversity, $\log
\left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$, $D_{i,t}$ is the percent
change in days fished, $\log
\left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and $I_{i,t}$ is an
indicator variable that takes a value $0$ if $\mathrm{div}_{i,t} \ge
\mathrm{div}_{i,t-1}$ (species diversity increasing; generalizing) and takes
a value $1$ if $\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$ (species diversity
decreasing; specializing). Estimating a separate coefficient for the effect of
increasing diversity and decreasing diversity allows the costs and benefits of
these changes to be asymmetric.

We use $\log(R_{i,t-1}$) as an offset, and thus our model describes the percent
change in revenue, i.e., $\log(R_{i,t}) - \log(R_{i,t-1})
= \log(R_{i,t}/R_{i,t-1})$. We model change in revenue because preliminary
analysis revealed non-stationary trends in revenue (resulting from experience or
other unmodeled factors). This approach eliminates having to deal with the
non-stationary time series and eliminates having to estimate thousands of
mean-revenue intercepts associated with individual fishers. We removed
person-year combinations without a subsequent year of revenue.

The $\beta$ terms represent estimated coefficients, some of which are allowed to
vary as "random effect" terms. The intercepts $\beta_{0,j,t}$ are allowed to
vary by strategy-year combinations, are centered on zero, and are constrained by
a normal distribution:

<!-- Ole: I'd use tau (or something) for the variances here so that people don't have to look at the subscripts to differentiate between variances at different hierarchical levels. -->

\begin{equation}
\beta_{0,j,t} \sim \mathrm{N} \left(0, \sigma^2_{\beta_{j,t}} \right).
\end{equation}

\noindent This approach, combined with the lack of a global intercept,
constrains the mean change in log revenue across all fishers and years to be
zero. I.e., we assume that the data are stationary after accounting for all
covariates and including an intercept that varies with strategy-year
combinations. Initial versions of the model confirmed that the global intercept
along with strategy-level intercepts would be estimated at almost exactly zero.

The slopes $\beta_{1,j}$ and $\beta_{1,j}$ are then allowed to vary by strategy
with estimated means:

\begin{align}
\beta_{1,j} &\sim \mathrm{N} \left(\mu_{\beta_1}, \sigma^2_{\beta_1}\right),\\
\beta_{2,j} &\sim \mathrm{N} \left(\mu_{\beta_2}, \sigma^2_{\beta_2}\right).
\end{align}

\noindent The $\beta_{1,j}$'s represent how much a 1\% decrease in species
diversity within a strategy translates to a Y\% change in revenue. The
$\beta_{2,j}$'s represent how much a 1\% increase in species diversity within
a strategy translates to a Y\% change in revenue. These are after controlling
for percent change in days fished, $\beta_3$, and the interaction between days
fished and percent change in species diversity, $\beta_4$ and $\beta_5$.

Whereas a typical hierarchical linear regression model would estimate a single
residual error scale term, $\sigma$, we model the scale of the residual error
with another hierarchical model with a similar form:

\begin{equation}
\label{eq:cv}
\widehat{\sigma_{i,t}} =
\exp \left( \gamma_{0,j} -
\gamma_{1,j} S_{i,t} I_{i,t} +
\gamma_{2,j} S_{i,t} I_{i,t} +
\gamma_3 D_{i,t} -
\gamma_4 S_{i,t} D_{i,t} I_{i,t} +
\gamma_5 S_{i,t} D_{i,t} I_{i,t}\right).
\end{equation}

\noindent Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively, percent change
in species diversity and an indicator variable for increasing or decreasing
species diversity. We exponentiate the equation to ensure all scale parameters,
$\sigma_{i,t}$, are positive. The $\gamma$'s are estimated coefficients, three
of which are allowed to vary by strategy, constrained by normal distributions:

\begin{align}
\gamma_{0,j}  &\sim \mathrm{N} \left(\eta_0 + \eta_1 M_j + \eta_2 E_j, \sigma^2_{\gamma_0}\right),\\
\gamma_{1,j}  &\sim \mathrm{N} \left(\mu_{\gamma_1}, \sigma^2_{\gamma_1}\right),\\
\gamma_{2,j}  &\sim \mathrm{N} \left(\mu_{\gamma_2}, \sigma^2_{\gamma_2}\right).
\end{align}

\noindent The mean of the $\gamma_{0,j}$ coefficients is modelled with two
strategy-level predictors. The symbols $M_j$ and $E_j$ represent the raw mean
species diversity and mean combined fishing season length (effort) for strategy
$j$. The coefficients $\eta_1$ and $\eta_2$ represent predictors in this
strategy-level regression estimating effects across strategies.

>>>>

<!DOCTYPE html>
<html lang="en" class=" is-copy-enabled emoji-size-boost is-u2f-enabled">
  <head prefix="og: http://ogp.me/ns# fb: http://ogp.me/ns/fb# object: http://ogp.me/ns/object# article: http://ogp.me/ns/article# profile: http://ogp.me/ns/profile#">
    <meta charset='utf-8'>

    <link crossorigin="anonymous" href="https://assets-cdn.github.com/assets/frameworks-c64b206fb6104fd10a7b4f7460f970885149c62649a7414df084683d7a0f7bf3.css" integrity="sha256-xksgb7YQT9EKe090YPlwiFFJxiZJp0FN8IRoPXoPe/M=" media="all" rel="stylesheet" />
    <link crossorigin="anonymous" href="https://assets-cdn.github.com/assets/github-10cfb952863ec274f56e591f95dd8cd5e2c2516671d7989a327f7bf753c08909.css" integrity="sha256-EM+5UoY+wnT1blkfld2M1eLCUWZx15iaMn9791PAiQk=" media="all" rel="stylesheet" />

    <meta http-equiv="X-UA-Compatible" content="IE=edge">
    <meta http-equiv="Content-Language" content="en">
    <meta name="viewport" content="width=device-width">

    <title>pfx-commercial/ms.Rmd at master · NCEAS/pfx-commercial</title>
    <link rel="search" type="application/opensearchdescription+xml" href="/opensearch.xml" title="GitHub">
    <link rel="fluid-icon" href="https://github.com/fluidicon.png" title="GitHub">
    <link rel="apple-touch-icon" href="/apple-touch-icon.png">
    <link rel="apple-touch-icon" sizes="57x57" href="/apple-touch-icon-57x57.png">
    <link rel="apple-touch-icon" sizes="60x60" href="/apple-touch-icon-60x60.png">
    <link rel="apple-touch-icon" sizes="72x72" href="/apple-touch-icon-72x72.png">
    <link rel="apple-touch-icon" sizes="76x76" href="/apple-touch-icon-76x76.png">
    <link rel="apple-touch-icon" sizes="114x114" href="/apple-touch-icon-114x114.png">
    <link rel="apple-touch-icon" sizes="120x120" href="/apple-touch-icon-120x120.png">
    <link rel="apple-touch-icon" sizes="144x144" href="/apple-touch-icon-144x144.png">
    <link rel="apple-touch-icon" sizes="152x152" href="/apple-touch-icon-152x152.png">
    <link rel="apple-touch-icon" sizes="180x180" href="/apple-touch-icon-180x180.png">
    <meta property="fb:app_id" content="1401488693436528">

      <meta content="https://avatars2.githubusercontent.com/u/2159348?v=3&amp;s=400" name="twitter:image:src" /><meta content="@github" name="twitter:site" /><meta content="summary" name="twitter:card" /><meta content="NCEAS/pfx-commercial" name="twitter:title" /><meta content="Contribute to pfx-commercial development by creating an account on GitHub." name="twitter:description" />
      <meta content="https://avatars2.githubusercontent.com/u/2159348?v=3&amp;s=400" property="og:image" /><meta content="GitHub" property="og:site_name" /><meta content="object" property="og:type" /><meta content="NCEAS/pfx-commercial" property="og:title" /><meta content="https://github.com/NCEAS/pfx-commercial" property="og:url" /><meta content="Contribute to pfx-commercial development by creating an account on GitHub." property="og:description" />
      <meta name="browser-stats-url" content="https://api.github.com/_private/browser/stats">
    <meta name="browser-errors-url" content="https://api.github.com/_private/browser/errors">
    <link rel="assets" href="https://assets-cdn.github.com/">
    <link rel="web-socket" href="wss://live.github.com/_sockets/NTA0Njg4NDo0OTE1ZGE4NjRjZjExYmU3YzgwMDFmMGEwN2M4ZTIxYzplM2U0Yjk4MWFjMjg5N2U4MTMyMDM2YzMxYWE3NzBmNzg0MmY0OTgwZmJhYTQ1Zjc2NzBhYzg4ZTdiNjM2MmI5--9f3c14941fd0bc15eb2fce42405037f6cb2a11f5">
    <meta name="pjax-timeout" content="1000">
    <link rel="sudo-modal" href="/sessions/sudo_modal">
    <meta name="request-id" content="1812FC12:15610:276B4F4D:58473282" data-pjax-transient>

    <meta name="msapplication-TileImage" content="/windows-tile.png">
    <meta name="msapplication-TileColor" content="#ffffff">
    <meta name="selected-link" value="repo_source" data-pjax-transient>

    <meta name="google-site-verification" content="KT5gs8h0wvaagLKAVWq8bbeNwnZZK1r1XQysX3xurLU">
<meta name="google-site-verification" content="ZzhVyEFwb7w3e0-uOTltm8Jsck2F5StVihD0exw2fsA">
    <meta name="google-analytics" content="UA-3769691-2">

<meta content="collector.githubapp.com" name="octolytics-host" /><meta content="github" name="octolytics-app-id" /><meta content="1812FC12:15610:276B4F4D:58473282" name="octolytics-dimension-request_id" /><meta content="5046884" name="octolytics-actor-id" /><meta content="eric-ward" name="octolytics-actor-login" /><meta content="d7c8c13b6921d5dbb2d08511e76e4d3922fb6f78877f69812f4e2cc5b901f5c9" name="octolytics-actor-hash" />
<meta content="/&lt;user-name&gt;/&lt;repo-name&gt;/blob/show" data-pjax-transient="true" name="analytics-location" />

  <meta class="js-ga-set" name="dimension1" content="Logged In">

        <meta name="hostname" content="github.com">
    <meta name="user-login" content="eric-ward">

        <meta name="expected-hostname" content="github.com">
      <meta name="js-proxy-site-detection-payload" content="OTZkMDAzYjQ5M2JjYTZlYzA5MmU0MTdkODA1ZDk5MWU1MWY3Y2RjNzhlMzdlN2FlNWZkZjUxZjc3ZDRhYmQ5Nnx7InJlbW90ZV9hZGRyZXNzIjoiMjQuMTguMjUyLjE4IiwicmVxdWVzdF9pZCI6IjE4MTJGQzEyOjE1NjEwOjI3NkI0RjREOjU4NDczMjgyIiwidGltZXN0YW1wIjoxNDgxMDYwOTk5LCJob3N0IjoiZ2l0aHViLmNvbSJ9">

      <link rel="mask-icon" href="https://assets-cdn.github.com/pinned-octocat.svg" color="#000000">
      <link rel="icon" type="image/x-icon" href="https://assets-cdn.github.com/favicon.ico">

    <meta name="html-safe-nonce" content="306849f02c6158107b6119c3ea451ae14657b020">
    <meta content="f2086a1d058512824537f14ace9f906f412f8160" name="form-nonce" />

    <meta http-equiv="x-pjax-version" content="77135ce72af7054f190f2da8838d6ffb">

  <meta name="description" content="Contribute to pfx-commercial development by creating an account on GitHub.">
  <meta name="go-import" content="github.com/NCEAS/pfx-commercial git https://github.com/NCEAS/pfx-commercial.git">

  <meta content="2159348" name="octolytics-dimension-user_id" /><meta content="NCEAS" name="octolytics-dimension-user_login" /><meta content="45410050" name="octolytics-dimension-repository_id" /><meta content="NCEAS/pfx-commercial" name="octolytics-dimension-repository_nwo" /><meta content="true" name="octolytics-dimension-repository_public" /><meta content="false" name="octolytics-dimension-repository_is_fork" /><meta content="45410050" name="octolytics-dimension-repository_network_root_id" /><meta content="NCEAS/pfx-commercial" name="octolytics-dimension-repository_network_root_nwo" />
  <link href="https://github.com/NCEAS/pfx-commercial/commits/master.atom" rel="alternate" title="Recent Commits to pfx-commercial:master" type="application/atom+xml">

      <link rel="canonical" href="https://github.com/NCEAS/pfx-commercial/blob/master/portfolio/text/ms.Rmd" data-pjax-transient>
  </head>

  <body class="logged-in  env-production macintosh vis-public page-blob">
    <div id="js-pjax-loader-bar" class="pjax-loader-bar"><div class="progress"></div></div>
    <a href="#start-of-content" tabindex="1" class="accessibility-aid js-skip-to-content">Skip to content</a>

        <div class="header header-logged-in true" role="banner">
  <div class="container clearfix">

    <a class="header-logo-invertocat" href="https://github.com/" data-hotkey="g d" aria-label="Homepage" data-ga-click="Header, go to dashboard, icon:logo">
  <svg aria-hidden="true" class="octicon octicon-mark-github" height="28" version="1.1" viewBox="0 0 16 16" width="28"><path fill-rule="evenodd" d="M8 0C3.58 0 0 3.58 0 8c0 3.54 2.29 6.53 5.47 7.59.4.07.55-.17.55-.38 0-.19-.01-.82-.01-1.49-2.01.37-2.53-.49-2.69-.94-.09-.23-.48-.94-.82-1.13-.28-.15-.68-.52-.01-.53.63-.01 1.08.58 1.23.82.72 1.21 1.87.87 2.33.66.07-.52.28-.87.51-1.07-1.78-.2-3.64-.89-3.64-3.95 0-.87.31-1.59.82-2.15-.08-.2-.36-1.02.08-2.12 0 0 .67-.21 2.2.82.64-.18 1.32-.27 2-.27.68 0 1.36.09 2 .27 1.53-1.04 2.2-.82 2.2-.82.44 1.1.16 1.92.08 2.12.51.56.82 1.27.82 2.15 0 3.07-1.87 3.75-3.65 3.95.29.25.54.73.54 1.48 0 1.07-.01 1.93-.01 2.2 0 .21.15.46.55.38A8.013 8.013 0 0 0 16 8c0-4.42-3.58-8-8-8z"/></svg>
</a>

        <div class="header-search scoped-search site-scoped-search js-site-search" role="search">
  <!-- '"` --><!-- </textarea></xmp> --></option></form><form accept-charset="UTF-8" action="/NCEAS/pfx-commercial/search" class="js-site-search-form" data-scoped-search-url="/NCEAS/pfx-commercial/search" data-unscoped-search-url="/search" method="get"><div style="margin:0;padding:0;display:inline"><input name="utf8" type="hidden" value="&#x2713;" /></div>
    <label class="form-control header-search-wrapper js-chromeless-input-container">
      <div class="header-search-scope">This repository</div>
      <input type="text"
        class="form-control header-search-input js-site-search-focus js-site-search-field is-clearable"
        data-hotkey="s"
        name="q"
        placeholder="Search"
        aria-label="Search this repository"
        data-unscoped-placeholder="Search GitHub"
        data-scoped-placeholder="Search"
        autocapitalize="off">
    </label>
</form></div>

      <ul class="header-nav float-left" role="navigation">
        <li class="header-nav-item">
          <a href="/pulls" aria-label="Pull requests you created" class="js-selected-navigation-item header-nav-link" data-ga-click="Header, click, Nav menu - item:pulls context:user" data-hotkey="g p" data-selected-links="/pulls /pulls/assigned /pulls/mentioned /pulls">
            Pull requests
</a>        </li>
        <li class="header-nav-item">
          <a href="/issues" aria-label="Issues you created" class="js-selected-navigation-item header-nav-link" data-ga-click="Header, click, Nav menu - item:issues context:user" data-hotkey="g i" data-selected-links="/issues /issues/assigned /issues/mentioned /issues">
            Issues
</a>        </li>
          <li class="header-nav-item">
            <a class="header-nav-link" href="https://gist.github.com/" data-ga-click="Header, go to gist, text:gist">Gist</a>
          </li>
      </ul>

<ul class="header-nav user-nav float-right" id="user-links">
  <li class="header-nav-item">

    <a href="/notifications" aria-label="You have unread notifications" class="header-nav-link notification-indicator tooltipped tooltipped-s js-socket-channel js-notification-indicator" data-channel="tenant:1:notification-changed:5046884" data-ga-click="Header, go to notifications, icon:unread" data-hotkey="g n">
        <span class="mail-status unread"></span>
        <svg aria-hidden="true" class="octicon octicon-bell" height="16" version="1.1" viewBox="0 0 14 16" width="14"><path fill-rule="evenodd" d="M14 12v1H0v-1l.73-.58c.77-.77.81-2.55 1.19-4.42C2.69 3.23 6 2 6 2c0-.55.45-1 1-1s1 .45 1 1c0 0 3.39 1.23 4.16 5 .38 1.88.42 3.66 1.19 4.42l.66.58H14zm-7 4c1.11 0 2-.89 2-2H5c0 1.11.89 2 2 2z"/></svg>
</a>
  </li>

  <li class="header-nav-item dropdown js-menu-container">
    <a class="header-nav-link tooltipped tooltipped-s js-menu-target" href="/new"
       aria-label="Create new…"
       data-ga-click="Header, create new, icon:add">
      <svg aria-hidden="true" class="octicon octicon-plus float-left" height="16" version="1.1" viewBox="0 0 12 16" width="12"><path fill-rule="evenodd" d="M12 9H7v5H5V9H0V7h5V2h2v5h5z"/></svg>
      <span class="dropdown-caret"></span>
    </a>

    <div class="dropdown-menu-content js-menu-content">
      <ul class="dropdown-menu dropdown-menu-sw">

<a class="dropdown-item" href="/new" data-ga-click="Header, create new repository">
  New repository
</a>

  <a class="dropdown-item" href="/new/import" data-ga-click="Header, import a repository">
    Import repository
  </a>

<a class="dropdown-item" href="https://gist.github.com/" data-ga-click="Header, create new gist">
  New gist
</a>

  <a class="dropdown-item" href="/organizations/new" data-ga-click="Header, create new organization">
    New organization
  </a>

  <div class="dropdown-divider"></div>
  <div class="dropdown-header">
    <span title="NCEAS/pfx-commercial">This repository</span>
  </div>
    <a class="dropdown-item" href="/NCEAS/pfx-commercial/issues/new" data-ga-click="Header, create new issue">
      New issue
    </a>

      </ul>
    </div>
  </li>

  <li class="header-nav-item dropdown js-menu-container">
    <a class="header-nav-link name tooltipped tooltipped-sw js-menu-target" href="/eric-ward"
       aria-label="View profile and more"
       data-ga-click="Header, show menu, icon:avatar">
      <img alt="@eric-ward" class="avatar" height="20" src="https://avatars2.githubusercontent.com/u/5046884?v=3&amp;s=40" width="20" />
      <span class="dropdown-caret"></span>
    </a>

    <div class="dropdown-menu-content js-menu-content">
      <div class="dropdown-menu dropdown-menu-sw">
        <div class="dropdown-header header-nav-current-user css-truncate">
          Signed in as <strong class="css-truncate-target">eric-ward</strong>
        </div>

        <div class="dropdown-divider"></div>

        <a class="dropdown-item" href="/eric-ward" data-ga-click="Header, go to profile, text:your profile">
          Your profile
        </a>
        <a class="dropdown-item" href="/eric-ward?tab=stars" data-ga-click="Header, go to starred repos, text:your stars">
          Your stars
        </a>
        <a class="dropdown-item" href="/explore" data-ga-click="Header, go to explore, text:explore">
          Explore
        </a>
          <a class="dropdown-item" href="/integrations" data-ga-click="Header, go to integrations, text:integrations">
            Integrations
          </a>
        <a class="dropdown-item" href="https://help.github.com" data-ga-click="Header, go to help, text:help">
          Help
        </a>

        <div class="dropdown-divider"></div>

        <a class="dropdown-item" href="/settings/profile" data-ga-click="Header, go to settings, icon:settings">
          Settings
        </a>

        <!-- '"` --><!-- </textarea></xmp> --></option></form><form accept-charset="UTF-8" action="/logout" class="logout-form" data-form-nonce="f2086a1d058512824537f14ace9f906f412f8160" method="post"><div style="margin:0;padding:0;display:inline"><input name="utf8" type="hidden" value="&#x2713;" /><input name="authenticity_token" type="hidden" value="PnbQhWTo21/c91t6711esJeC3XADkQpt6F5tc07PrUW8XnSQgyd36jUpfGEHM3UbQRUVuX3tNIMOckYBS+V6lg==" /></div>
          <button type="submit" class="dropdown-item dropdown-signout" data-ga-click="Header, sign out, icon:logout">
            Sign out
          </button>
</form>      </div>
    </div>
  </li>
</ul>

  </div>
</div>

    <div id="start-of-content" class="accessibility-aid"></div>

      <div id="js-flash-container">
</div>

    <div role="main">
        <div itemscope itemtype="http://schema.org/SoftwareSourceCode">
    <div id="js-repo-pjax-container" data-pjax-container>

<div class="pagehead repohead instapaper_ignore readability-menu experiment-repo-nav">
  <div class="container repohead-details-container">

<ul class="pagehead-actions">

  <li>
        <!-- '"` --><!-- </textarea></xmp> --></option></form><form accept-charset="UTF-8" action="/notifications/subscribe" class="js-social-container" data-autosubmit="true" data-form-nonce="f2086a1d058512824537f14ace9f906f412f8160" data-remote="true" method="post"><div style="margin:0;padding:0;display:inline"><input name="utf8" type="hidden" value="&#x2713;" /><input name="authenticity_token" type="hidden" value="hFCZaDumpg36s+UNPQAtiT2CABWrtFfxVyHcBMJBBc0GeD193GkKuBNtwhbVbgYi6xXI3NXIaR+xDfd2x2vSHg==" /></div>      <input class="form-control" id="repository_id" name="repository_id" type="hidden" value="45410050" />

        <div class="select-menu js-menu-container js-select-menu">
          <a href="/NCEAS/pfx-commercial/subscription"
            class="btn btn-sm btn-with-count select-menu-button js-menu-target" role="button" tabindex="0" aria-haspopup="true"
            data-ga-click="Repository, click Watch settings, action:blob#show">
            <span class="js-select-button">
              <svg aria-hidden="true" class="octicon octicon-eye" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M8.06 2C3 2 0 8 0 8s3 6 8.06 6C13 14 16 8 16 8s-3-6-7.94-6zM8 12c-2.2 0-4-1.78-4-4 0-2.2 1.8-4 4-4 2.22 0 4 1.8 4 4 0 2.22-1.78 4-4 4zm2-4c0 1.11-.89 2-2 2-1.11 0-2-.89-2-2 0-1.11.89-2 2-2 1.11 0 2 .89 2 2z"/></svg>
              Unwatch
            </span>
          </a>
          <a class="social-count js-social-count"
            href="/NCEAS/pfx-commercial/watchers"
            aria-label="18 users are watching this repository">
            18
          </a>
